---
title: "STATS183 Project 2"
author: "Takao Oba"
date: "2023-04-14"
output:
  pdf_document:
    latex_engine: xelatex
---


## a. Refer to the lecture material and the paper “An Analytic Derivation of the Efficient Portfolio Frontier,” (JFQA, Robert Merton, 1972).Compute A, B, C, D.

```{r}
#Read your csv file:
a <- read.csv("/Users/takaooba/STATS 183/stockData.csv", sep=",", header=TRUE)
train <- a[1:60,]
test <- a[61:dim(a)[1],]

#Convert adjusted close prices into returns:
r <- (train[-1,3:ncol(train)]-train[-nrow(train),3:ncol(train)])/train[-nrow(train),3:ncol(train)]

#Compute mean vector:
means <- colMeans(r[-1]) #Without ^GSPC

#Compute variance covariance matrix:
covmat <- cov(r[-1]) #Without ^GSPC

#Compute the vector of variances:
variances <- diag(covmat)

#Compute the vector of standard deviations:
stdev <- diag(covmat)^.5

# mean vector of SP500
means_sp500 <- mean(r[,1])
stdev_sp500 <- sd(r[,1])

# one vector
ones <- rep(1,30)
```

Computing A,B,CD
```{r}
#Compute A:
A <- t(ones) %*% solve(covmat) %*% means
A

#Compute B:
B <- t(means) %*% solve(covmat) %*% means
B

#Compute C:
C <- t(ones) %*% solve(covmat) %*% ones
C

#Compute D:
D <- B*C - A^2
D
```

## b. Compute the values of λ1 and λ2 (the two Lagrange multipliers).

```{r}
# Let E = 0.15
E = 0.15

# λ1
lambda1 <- (C*E-A)/D
lambda1

# λ2
lambda2 <- (B-A*E)/D
lambda2
```


## c. Suppose an investor has a prescribed expected return E. Find the composition of the efficient portfolio given the return E. Note: You need to choose a value of E.

```{r}
x2=as.numeric(lambda1)*solve(covmat) %*% means + 
as.numeric(lambda2)* solve(covmat) %*% ones
x2

X2 <- solve(covmat) %*% (lambda1*means + lambda2*ones)
return2 <- t(X2) %*% means
var2 <- t(X2) %*% covmat %*% X2
sd2 <- sqrt(var2)
```


## d. Use your data to plot the frontier in the mean-variance space (parabola)

```{r}
#Give values for E:
E <- seq(-5,5,.1)

#Compute sigma2 as a function of A,B,C,D, and E:
sigma2 <- (C*E^2 - 2*A*E +B) /D

#Or plot E against sigma2:
plot(sigma2, E,type="l", xlab=expression(sigma^2), xlim = c(0, 1),   ylim = c(-2,2), main = "Frontier in the mean-variance space")

#Add the minimum risk portfolio:
points(1/C, A/C, pch=19)

```


<!-- ```{r} -->
<!-- plot(0, A/C, main = "Portfolio possibilities curve", xlab = "Risk (standard deviation)", -->
<!--   ylab = "Expected Return", type = "n", -->
<!--   xlim = c(-2*sqrt(1/C), 4*sqrt(1/C)),  -->
<!--   ylim = c(-2*A/C, 4*A/C)) -->

<!-- #Plot center of the hyperbola: -->
<!--     points(0, A/C, pch = 19) -->

<!-- #Plot transverse and conjugate axes: -->
<!--     abline(v = 0, lwd = 2) #Also this is the y-axis. -->
<!--     abline(h = A/C) -->

<!-- #Plot the x-axis: -->
<!--     abline(h = 0, lwd = 2) -->

<!-- #Plot the minimum risk portfolio: -->
<!--     points(sqrt(1/C), A/C, pch=19) -->

<!-- #Efficient frontier: -->
<!--     minvar <- 1/C -->
<!--     minE <- A/C -->
<!--     sdeff <- seq((minvar)^0.5, 1, by = 0.0001) -->
<!--     options(warn = -1) -->
<!--     y1 <- (A + sqrt(D*(C*sdeff^2 - 1)))*(1/C)  -->
<!--     y2 <- (A - sqrt(D*(C*sdeff^2 - 1)))*(1/C)  -->
<!--     options(warn = 0) -->

<!--     points(sdeff, y1, type = "l", col = "red") -->
<!--     points(sdeff, y2, type = "l") -->
<!-- ``` -->



## e. Use your data to plot the frontier in the mean-standard deviation space using the hyperbola method.

```{r}
#Hyperbola:
#Efficient frontier:
    minvar <- 1/C
    minE <- A/C
    sdeff <- seq((minvar)^0.5, 1, by = 0.0001)
    
#    options(warn = -1)
    y1 <- (A + sqrt(D*(C*sdeff^2 - 1)))*(1/C) 
    y2 <- (A - sqrt(D*(C*sdeff^2 - 1)))*(1/C) 
#    options(warn = 0)

plot(sdeff, y1, type = "n",xlim=c(0 ,0.15), ylim=c(-0.15,0.2), 
     xlab="Portfolio standard deviation", ylab="Expected return", 
     xaxt="no", yaxt="no", main = "Frontier in the Mean-SD Space")

axis(1, at=seq(0, 0.15, 0.02))
axis(2, at=seq(-0.15,0.2, 0.02))

    points(sdeff, y1, lwd=5,type = "l", col = "red")
    points(sdeff, y2, lwd=5,type = "l")

# min risk portfolio
points(sqrt(1/C), A/C, pch = 19, col = 10, lwd = 5)
```


## f. On the plot in (e) add the 30 stocks, the S&P500, the equal allocation portfolio, the minimum risk portfolio, and the portfolio in (c).

```{r}
# equal allocation
eq = rep(1/30, 30)

# mean of this portfolio
r_eq = t(eq) %*% means

# sd of this portfolio
sd_eq = sqrt(t(eq) %*% covmat %*% eq)
```



```{r}
plot(sdeff, y1, type = "n",xlim=c(0 ,0.15), ylim=c(-0.15,0.2), 
     xlab="Portfolio standard deviation", ylab="Expected return", 
     xaxt="no", yaxt="no", main = "Frontier in the Mean-SD Space")

axis(1, at=seq(0, 0.15, 0.02))
axis(2, at=seq(-0.15,0.2, 0.02))

    points(sdeff, y1, lwd=5,type = "l", col = "red")
    points(sdeff, y2, lwd=5,type = "l")

# min risk portfolio
points(sqrt(1/C), A/C, pch = 19, col = 10, lwd = 5)
# 30 stocks
points(stdev, means, pch = 19, col = 12)
# sp500
points(stdev_sp500, stdev_sp500, pch = 19, col = 3)
# equal allocation
points(sd_eq, r_eq, pch = 19, col = 7)
# part c
points(sd2, return2, pch = 19, col = 5)

legend(x = 'topleft',cex  = 0.65,  
       legend = c("Minimum Risk Portfolio","30 stocks", "sp500", "equal allocation", "E = 0.15"),
       text.col = c(10, 12, 3, 7, 5))

```


## g. Add three arbitrary portfolios on the plot of (c). You can choose any 30 weights with 
$$\Sigma_{i=1}^{30} x_i = 1$$
Three arbitrary portfolios
```{r}
# Portfolio 1, E = 0.03
E <- 0.03
lambda1_1 <- (C*E-A)/D
lambda2_1 <- (B-A*E)/D

X2_1 <- solve(covmat) %*% (lambda1_1*means + lambda2_1*ones)
return2_1 <- t(X2_1) %*% means
var2_1 <- t(X2_1) %*% covmat %*% X2_1
sd2_1 <- sqrt(var2_1)

# Portfolio 2, E = 0.08
E <- 0.08
lambda1_2 <- (C*E-A)/D
lambda2_2 <- (B-A*E)/D

X2_2 <- solve(covmat) %*% (lambda1_2*means + lambda2_2*ones)
return2_2 <- t(X2_2) %*% means
var2_2 <- t(X2_2) %*% covmat %*% X2_2
sd2_2 <- sqrt(var2_2)

# Portfolio 3, E = 0.12
E <- 0.12
lambda1_3 <- (C*E-A)/D
lambda2_3 <- (B-A*E)/D

X2_3 <- solve(covmat) %*% (lambda1_3*means + lambda2_3*ones)
return2_3 <- t(X2_3) %*% means
var2_3 <- t(X2_3) %*% covmat %*% X2_3
sd2_3 <- sqrt(var2_3)
```




```{r}
plot(sdeff, y1, type = "n",xlim=c(0 ,0.15), ylim=c(-0.15,0.2), 
     xlab="Portfolio standard deviation", ylab="Expected return", 
     xaxt="no", yaxt="no", main = "Frontier in the Mean-SD Space with 3 Portfolios Added")

axis(1, at=seq(0, 0.15, 0.02))
axis(2, at=seq(-0.15,0.2, 0.02))

    points(sdeff, y1, lwd=5,type = "l", col = "red")
    points(sdeff, y2, lwd=5,type = "l")

# min risk portfolio
points(sqrt(1/C), A/C, pch = 19, col = 10, lwd = 5)
# 30 stocks
points(stdev, means, pch = 19, col = 12)
# sp500
points(stdev_sp500, stdev_sp500, pch = 19, col = 3)
# equal allocation
points(sd_eq, r_eq, pch = 19, col = 7)
# part c
points(sd2, return2, pch = 19, col = 5)

points(sd2_1, return2_1, pch = 19, col = "black")
points(sd2_2, return2_2, pch = 19, col = "black")
points(sd2_3, return2_3, pch = 19, col = "black")

legend(x = 'topleft',cex  = 0.65,  
       legend = c("Minimum Risk Portfolio","30 stocks", "sp500", "equal allocation", "E = 0.15", "Added Portfolios"),
       text.col = c(10, 12, 3, 7, 5, "black"))

```


