---
title: "STATS183 Project 5"
author: "Takao Oba"
date: "2023-05-01"
output:
  pdf_document:
    latex_engine: xelatex
---

## Please answer the following questions assuming the single index model holds:

### a. Assume the single index model holds. Use only the stocks with positive betas in your data. Choose a value of Rf and find the optimal portfolio (point of tangency) using the optimization procedure as discussed in handout #12: http://www.stat.ucla.edu/~nchristo/statistics_c183_c283/statc183c283_tangent.pdf . The approach here is based on Z = Σ−1R.

```{r}
#Read your csv file:
a <- read.csv("/Users/takaooba/STATS 183/stockData.csv", sep=",", header=TRUE)
train <- a[1:60,]
test <- a[61:dim(a)[1],]

#Convert adjusted close prices into returns:
r <- (train[-1,3:ncol(train)]-train[-nrow(train),3:ncol(train)])/train[-nrow(train),3:ncol(train)]
```

```{r}
rr <- r[,-1]
#Compute the variance-covariance matrix:
var_covar <- cov(rr)
# var_covar

#Compute the inverse of the variance-covariance matrix:
var_covar_inv <- solve(var_covar)
# var_covar_inv

#Create the vector R:
Rf <- 0.002
R_ibar <- apply(rr, 2, mean)
R <- R_ibar-Rf
# R

#Compute the vector Z:
z <- var_covar_inv %*% R
# z

#Compute the vector X:
x <- z/sum(z)
# x

#Compute the expected return of portfolio G:
R_Gbar <- t(x) %*% R_ibar
R_Gbar

#Compute the variance and standard deviation of portfolio G:
var_G <- t(x) %*% var_covar %*% x
# var_G

sd_G <- var_G^0.5
sd_G

#Compute the slope:
slope <- (R_Gbar-Rf)/(sd_G)
# slope
```


Side note: using C*
```{r}
#Compute the betas:
covmat <- var(r)
beta <- covmat[1,-1] / covmat[1,1]

#Keep only the stocks with positive betas:
rrr <- r[,-c(1,which(beta<0)+1)]  

#all of the stocks have positive betas
length(rrr)
#Note: which(beta<0) gives the element in the beta vector with negative beta and add 1 because 
#the first column in the iitial data set is the index.  
# We also remove column 1 (index) from the initial data #set.

#Initialize
beta <- rep(0,ncol(rrr))
alpha <- rep(0,ncol(rrr))
mse <- rep(0,ncol(rrr))
Ribar <- rep(0,ncol(rrr))
Ratio <- rep(0,ncol(rrr))
stock <- rep(0,ncol(rrr))

#Risk free asset:
rf <- 0.005


#This for loop computes the required inputs:
for(i in 1:ncol(rrr)){
	q <- lm(data=rrr, formula=rrr[,i] ~ r[,1])
	beta[i] <- q$coefficients[2] 
	alpha[i] <- q$coefficients[1] 
     mse[i] <- summary(q)$sigma^2
	Ribar[i] <- q$coefficients[1]+q$coefficients[2]*mean(r[,1])
	Ratio[i] <- (Ribar[i]-rf)/beta[i] # (R_i - R_f)/B_i
    stock[i] <- i 
}


#So far we have this table:
xx <- (cbind(stock,alpha, beta, Ribar, mse, Ratio))




#Order the table based on the excess return to beta ratio:
A <- xx[order(-xx[,6]),]

col1 <- rep(0,nrow(A))
col2 <- rep(0,nrow(A))
col3 <- rep(0,nrow(A))
col4 <- rep(0,nrow(A))
col5 <- rep(0,nrow(A))


#Create the last 5 columns of the table:
col1 <- (A[,4]-rf)*A[,3]/A[,5]
col3 <- A[,3]^2/A[,5]
for(i in(1:nrow(A))) {
col2[i] <- sum(col1[1:i])
col4[i] <- sum(col3[1:i])
}

#So far we have:
# cbind(A, col1, col2, col3, col4)
```


Recall that we can find c* by the following

$$C^* = \frac{\sigma_m^2\:\Sigma\frac{\bar{R_i} - R_f}{\sigma_{\epsilon_i}^2} b_i}{1+\sigma_m^2\Sigma\frac{b_i^2}{\sigma_{\epsilon_i}^2}}$$

Where 
$$\Sigma\frac{\bar{R_i} - R_f}{\sigma_{\epsilon_i}^2} b_i$$
is represented in column2

and 
$$\Sigma\frac{b_i^2}{\sigma_{\epsilon_i}^2}$$
is represented in column 4

```{r}

#Compute the Ci (col5):
for(i in (1:nrow(A))) {
col5[i] <- var(r[,1])*col2[i]/(1+var(r[,1])*col4[i])
}


```


Since we have that short selling is allowed, we have
```{r}
#SHORT SALES ALLOWED:
#Compute the Zi:
z_short <- (A[,3]/A[,5])*(A[,6]-col5[nrow(A)])
#Compute the xi:
x_short <- z_short/sum(z_short)

weights <- data.frame("Stock Name" = colnames(covmat)[-1], 
                  "Weights" = x_short)
weights
                  

#The final table when short sales allowed:
Weights_with_short <- cbind(A, col1, col2, col3, col4, col5, z_short, x_short)
```


### b. Adjusting the betas: Adjust the betas using Blume’s and Vasicek’s techniques. For the Blume technique use the two periods: 01- Jan-2015 to 01-Jan-2020 and 01-Jan-2020 to 31-Mar-2023. For the Vasicek technique use only the period 01-Jan-2014 to 01-Jan-2019. Note: For the Blume technique our goal is to adjust the betas in 01-Jan-2020 to 31-Mar-2023 to be better forecasts for the betas in period 01-Apr-2023 to 01-Apr-2027. For the Vasicek technique our goal is to adjust the betas in 01-Jan-2015 to 01-Jan-2020 to be better forecasts for the betas in period 01-Jan-2020 to 31-Mar-2023.


Blume's Technique
We will be using two periods from 01-Jan-2015 to 01-Jan-2020 and 01-Jan-2020 to 31-Mar-2023.
```{r}
#Convert adjusted close prices into returns:
r1 <- (train[-1,3:ncol(train)]-train[-nrow(train),3:ncol(train)])/train[-nrow(train),3:ncol(train)]

r2 <- (test[-1,3:ncol(test)]-test[-nrow(test),3:ncol(test)])/test[-nrow(test),3:ncol(test)]

#Compute the variance covariance matrix of the returns for each period:
covmat1 <- var(r1)
covmat2 <- var(r2)

#Compute the betas in each period:
beta1 <- covmat1[1,-1] / covmat1[1,1]
beta2 <- covmat2[1,-1] / covmat2[1,1]

#Correlation between the betas in the two periods:
cor(beta1, beta2)


#Adjust betas using the Blume's technique:
q1 <- lm(beta2 ~ beta1)


beta3adj_blume <- q1$coef[1] + q1$coef[2]*beta2
beta3adj_blume
```

Vasicek's Technique
Use period 01-Jan-2015 to 01-Jan-2020
```{r}
#Vasicek's method:
beta2 <- rep(0,30)

alpha2 <- rep(0,30)

sigma_e2 <- rep(0,30)

var_beta2 <- rep(0,30)

for(i in 1:30){
	q <- lm(data=r1, formula=r1[,i+1] ~ r1[,1])
	beta2[i] <- q$coefficients[2] 
	alpha2[i] <- q$coefficients[1] 
	sigma_e2[i] <- summary(q)$sigma^2
	var_beta2[i] <- vcov(q)[2,2] 
}


#Adjusting the betas using the Vasicek's technique:
beta3adj_vasicek <- var_beta2*mean(beta2)/(var(beta2)+var_beta2) + 
var(beta2)*beta2/(var(beta2)+var_beta2)
beta3adj_vasicek
```


### c. Compute PRESS only for the Vasicek technique. (You can compute the PRESS only for the Vasicek technique because you have the actual betas in the period 01-Jan-2020 to 31-Mar-2023.)

Recall there are three components to the MSE:
$$MSE  = (\bar{A} - \bar{P})^2 + (1-\hat{b_1})^2S_p^2 + (1-R^2)S_A^2$$

Where the Bias term is
$$ (\bar{A} - \bar{P})^2$$
and the infefficiency term is
$$(1-\hat{b_1})^2S_p^2$$
and the random error component is 
$$(1-R^2)S_A^2$$
```{r}
# Actual betas from testing is as follows:
r2 <- (test[-1,3:ncol(test)]-test[-nrow(test),3:ncol(test)])/test[-nrow(test),3:ncol(test)]
beta2 <- covmat2[1,-1] / covmat2[1,1]

#Using Vasicek's technique:
#1.  Bias component:
V1 <- ( mean(beta2) - mean(beta3adj_vasicek) )^2

#2.  Inefficiency component:
q3 <- lm(beta2 ~ beta3adj_vasicek)
Sp32 <- (29/30)*var(beta3adj_vasicek)
V2 <- (1-q3$coef[2])^2*Sp32

#3.  Random error component:
Sa2 <- (29/30)*var(beta2)
rap32 <- ( cor(beta3adj_vasicek,beta2) )^2
V3 <- (1-rap32)*Sa2

V1+V2+V3
```
Recall we can also achieve
```{r}
PRESS <- sum((beta2 - beta3adj_vasicek)^2)/30
PRESS
```

