---
title: "stats183_hw9"
author: "Takao"
date: "2023-06-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Suppose a stock has annual expected return and standard deviation μ = 0.20 and σ = 0.25. The current price of the stock is s = $50. Suppose that ∆t = 1week.
a. Find the distribution of the return of the stock during ∆t.
b. Simulate the path of the stock from now until 1 year from now (52 weeks). Submit the random samples and the plot of the price of the stock against time.

```{r}
epsilon <- c(0,rnorm(52))
S <- c(50,rep(0,52))
DS <- rep(0,53)

for(i in(1:52)) {
          
         DS[i+1] <- 0.2*(1/52)*S[i] + 0.25*(sqrt(1/52))*S[i]*epsilon[i+1] 

         S[i+1] = S[i] + DS[i+1]
}

x <- seq(0,52)
xx <- as.data.frame(cbind(x, epsilon, DS, S))

plot(x, S, type="l", xlab="Periods", ylab="Stock price")

points(x,S)
```



Exercise 5:
Using the lognormal property of stock prices estimate the annual volatility of APPLE (ticker is AAPL) using the adjusted daily close prices for the period 01-March-2023 to 26-May-2023. Save the data in a csv file and then read the data in R as follows:
s1 <- read.csv("AAPL.csv", sep=",", header=TRUE)

```{r}
s1 <- read.csv("AAPL.csv", sep=",", header=TRUE)

numerator <- s1[,3]
denom <- c(NA, numerator[1:60])
frac <- numerator / denom
u_i <- log(frac)

df <- data.frame(cbind(s1, numerator, denom, frac, u_i))
head(df)


# using formula, we can calculate the sd of u_i
sum_u_i <- sum(u_i[-1]^2)
n_ubar <- length(u_i[-1])*((mean(u_i[-1]))^2)
var_aapl <- (1/(length(u_i[-1])-1)) * (sum_u_i - n_ubar)

volatility <- sqrt(var_aapl) * sqrt(252)
volatility
```






